---
title: "CFM_shiny"
vignette: >
  %\VignetteIndexEntry{CFM_shiny}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: '#>'
---

```{r}
#| label: setup
library(MSCC)
```

# 1. Introduction

`shiny_vis_cfm()` is an interactive Shiny viewer for `CFM_data` objects (predicted / annotated spectra from CFM-ID).

Layout:

| panel            | content                                         |
|------------------|-------------------------------------------------|
| Multi-CE spectra | plotly spectra (energy0/1/2 → CE 10/20/40)      |
| Precursor        | precursor molecule (`vis_smiles`)               |
| Fragment         | selected fragment molecule                      |
| Info             | precursor / fragment (id, SMILES, formula, m/z) |

Click a peak in the spectra to update Fragment and Info.

# 2. Usage

## 2.1 Create a CFM demo

`CFM_data` holds:

- `@peak_assignment` — peaks per CE with optional `fragment_id` / `fragment_score`
- `@fragment_define` — fragment SMILES and `fragment_mz`
- `@fragment_transition` — fragment graph edges (when from annotate)

Build one from a SMILES string via CFM-ID (Docker required):

```{r}
#| label: cfm-demo
#| eval: false
cfm <- get_CFM_data_from_smiles(
  smiles = "NCC(O)=O",
  compound_id = "glycine",
  adduct = "[M+H]+",
  check_cache = TRUE,
  cache_dir = tempdir()
)
cfm
```

Small molecules such as glycine (`NCC(O)=O`) are useful for a quick demo.

## 2.2 CFM shiny

Launch the viewer:

```{r}
#| label: shiny-vis
#| eval: false
shiny_vis_cfm(cfm)
```

Optionally plot the multi-CE spectra alone:

```{r}
#| label: plotly-only
#| eval: false
plotly_CFM_spectra(cfm)
```
