---
title: "Get started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(MSCC)
```

# Installation

Install the development version from GitHub:

```r
# install.packages("pak")
pak::pak("drruili/MSCC")
```

Or with `remotes`:

```r
# install.packages("remotes")
remotes::install_github("drruili/MSCC")
```

# Formulas, mass, and m/z

Parse a chemical formula to an element matrix, then compute exact mass or charged m/z:

```{r formulas}
chemform_parse("C6H12O6")
chemform_mz("C6H12O6")
chemform_mz("C6H12O6", charge = 1)
```

Sum or multiply formulas when building larger species:

```{r formula-arith}
chemform_sum("C6H12O6", "H2O")
chemform_multi("CH2", 3)
```

# Adducts and isotopes

Prefer adduct strings like `[M+H]+` / `[M-H]-`. Isotope labels use bracket notation (`[13]C`, `[2]H`, …):

```{r adduct-iso}
chemform_adduct("C6H12O6", "[M+H]+", value = "chemform")
chemform_isotope_label("C6H12O6", "[13]C", 3)
```

Optional enviPat isotope patterns (requires the **enviPat** package):

```{r envipat, eval = requireNamespace("enviPat", quietly = TRUE)}
chemform_isotopes_pattern_enviPat("C6H12O6")
```

# Formula candidates from accurate mass

Decompose a neutral exact mass or an ion m/z into candidate formulas. Optionally apply Seven Golden Rules filters:

```{r decompose}
chemform_decompose_mass(180.0634, ppm = 5)
chemform_decompose_mz(181.0707, charge = 1, ppm = 5)

chemform_check_seven_golden_rules(c("C6H12O6", "CH6N2"))
```

See the [Seven Golden Rules](articles/Seven_Golden_Rules.html) article for rule details.

# Molecules from SMILES

Build a `Molecule_igraph` (SDF + igraph + optional isotopomers). Default demo molecule is glycine:

```{r molecule, eval = FALSE}
mol <- get_Molecule_igraph_from_smiles("NCC(O)=O", id = "glycine")
mol@molecule_info
```

# Conventions

| Topic | Convention |
|-------|------------|
| Isotope notation | `[13]C`, `[2]H`, `[15]N`, `[18]O`, `[34]S` |
| Adduct strings | Prefer `[M+H]+` / `[M-H]-` |
| Polarity | `0` = negative, `1` = positive |

# Next steps

- Reference index: function-level help on the package site
- Articles: formula decomposition, Seven Golden Rules, RDKit helpers
